### 2011-08-01 docs/pubs/EIM-2011/application.tex
finished edits for final version

The materialization routine used for semantic data queries laid the
groundwork for enabling data integration. In addition to inspecting
the data for values within a range and returning the data sets that
contain a match, the data integration feature of the Semtools Web
application goes further by constructing a synthetic data product
(table) that represents the complete results of the query in terms of
both the attributes and the values that are returned. Each original
data set may have very different syntactic structures (e.g. column
number, naming, order) but could share a subset of attributes that are
semantically compatible as defined in accompanying annotations. These
compatible, in-common attributes become the facets and metrics in a
synthetic data set; their values filtered
accordingly. \figref{fig:integration} further illustrates the data
integration support provided in the current implementation of the Web
application. Consider two data packages, denoted by A and B in
\figref{fig:integration}.  Annotations (denoted C and D) are used to
determine semantically equivalent data attributes contained in the
data sets (denoted by E and F). The attributes {\tt plot} and {\tt
  site} are considered equivalent measurements of the characteristic
Location; attributes {\tt weight} and {\tt wt} both map to the same
characteristic Mass. The Semantic Mediation API parses each annotation
and then computes an equivalence mapping among attributes based on
their corresponding annotations. The Data Manager Library is then used
to load the data sets and then query each data set to produce and
merge the final, synthetic result data set.

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{images/integration.pdf}
\caption{Data integration query across multiple data packages (A, B).
  Annotations (C, D) are used to determine semantically equivalent
  data attributes contained in the data objects (E, F). Attributes
  {\tt plot} and {\tt site} are considered equivalent measurements of the
  characteristic Location; attributes {\tt weight} and {\tt wt} both map to
  the same characteristic Mass. The Semantic Mediation API utilizes
  the Data Manager Library to load and query the source data informed
  by semantic similarities between the structurally disparate data
  attributes.}
\label{fig:integration}
\end{figure}

>>>>

The materialization routine for semantic data queries can help in
enabling data integration. In addition to inspecting data for values
within a range and returning the data sets that contain a match, the
data integration feature of the Semtools Web application goes further
by constructing a synthetic data product (table) that represents the
complete results of the query in terms of both the attributes and the
values that are returned. Each original data set may have very
different syntactic structures (e.g., column number, naming, order)
but could share a subset of attributes that are semantically
compatible as defined in accompanying annotations. These compatible
attributes can then become the basis for a synthetic data set.
\figref{fig:integration} illustrates the data integration support
provided in the current implementation of the Web
application. Consider the two data packages (denoted A and B) in
\figref{fig:integration}.  Annotations (denoted C and D) are used to
determine semantically equivalent data attributes contained in the
data sets (denoted by E and F). The attributes {\tt plot} and {\tt
  site} are considered equivalent measurements of the characteristic
Location; attributes {\tt weight} and {\tt wt} both map to the same
characteristic Mass. The Semantic Mediation API computes an
equivalence among attributes based on their corresponding
annotations. The Data Manager Library is then used to load the data
sets and then query each data set to produce and merge a synthetic
result data set.

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{images/integration.pdf}
\caption{Integration query across multiple data packages (A, B).
  Annotations (C, D) determine semantically equivalent data attributes
  contained in the data objects (E, F). Attributes {\tt plot} and {\tt
    site} are considered equivalent measurements of the characteristic
  Location; attributes {\tt weight} and {\tt wt} both map to the same
  characteristic Mass. The Semantic Mediation API utilizes the Data
  Manager Library to load and query the source data informed by
  semantic similarities between the structurally disparate data
  attributes.}
\label{fig:integration}
\end{figure}

### 2011-08-01 docs/pubs/EIM-2011/framework.tex
finished edits for final version

The Semtools project has focused development efforts on three main
components: a Java library for accessing and manipulating OBOE
ontology extensions and semantic annotations, an annotation plugin for
the Morpho metadata editor, and query extensions for the Metacat data
catalog. Below we describe these components and give a brief
overview of the OBOE model as well as the semantic annotation approach
used in Semtools. For a more in-depth presentation of OBOE see
\cite{madin07:_ontol_for_descr_and_synth,bowers08}.

>>>>

The Semtools project has focused development efforts on three main
components: a Java library to access and manipulate OBOE ontologies
and semantic annotations, an annotation plugin for the Morpho metadata
editor, and query extensions for the Metacat data catalog. Below we
briefly describe the OBOE model, the semantic annotation approach used
by Semtools, and the corresponding software components. For a more
in-depth presentation of OBOE see
\cite{madin07:_ontol_for_descr_and_synth,bowers08}.
